---
title: "Metaweb of European vertebrates analysis"
author: "Marc Ohlmann"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{vertebrates}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  screenshot.force = FALSE, 
  echo = TRUE,
  rows.print = 5,
  message = FALSE, 
  warning = FALSE)
```

## The European vertebrates data set

Metaweb of potential interactions between European terrestrial vertebrates. This data set is extracted from Maiorano et al. 2020 and O'Connor et al. 2020. It contains $1122$ vertebrate species (birds, mammals, amphibians and reptiles) and $49883$ expert-knowledge potential interactions. O'Connor et al. 2020 applied a Stochastic Block Model (SBM), a method that group together nodes with similar edge probability pattern (see Daudin et al.). This method inferred $46$ species groups. We also have classification of species in $4$ classes (Amphibians, Birds, Mammals and Reptiles).

In this vignette, we aim at representing this large network using SBM groups.

## Loading the data set

```{r packages}
library(metanetwork)
library(igraph)
library(ggplot2)
```

```{r data}
data("meta_vrtb")
print(meta_vrtb)
getwd()
```

### Append aggregated networks

To compute aggregated networks at the SBM group level and Class level, we use the `append_agg_nets` method.

```{r append agg nets}
meta_vrtb = append_agg_nets(meta_vrtb)
```

### Compute trophic levels

In order to represent this metaweb, we compute trophic levels since it is the first axis of 'metanetwork' layout.

```{r TL}
meta_vrtb = compute_TL(meta_vrtb)
```

We represent the distribution of trophic levels of the $4$ classes and $46$ SBM groups.

```{r distrib TL,fig.height = 7, fig.width = 7}
TL_df = cbind(name = V(meta_vrtb$metaweb)$name,
              TL = V(meta_vrtb$metaweb)$TL,
              group = meta_vrtb$trophicTable[V(meta_vrtb$metaweb),"group"],
              class = meta_vrtb$trophicTable[V(meta_vrtb$metaweb),"Class"]) %>%
        as.data.frame()

TL_df$TL = as.numeric(TL_df$TL)

ggplot(TL_df, aes(x=class, y=TL)) +
  geom_boxplot() + theme_classic()

ggplot(TL_df, aes(x=reorder(group, TL), y=TL)) +
  geom_boxplot() + theme_classic() + xlab("group") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

We see that the four classes contain species of various trophic levels. SBM groups are more ordered in terms of trophic level, even if several trophic groups have similar trophic levels.

```{r}
#this group contains several eagles species
which(meta_vrtb$trophicTable[,"group"] == 17) %>% names()
#single group made of eagle-owl only
which(meta_vrtb$trophicTable[,"group"] == 46) %>% names()
#this one contain wolf and fox
which(meta_vrtb$trophicTable[,"group"] == 42) %>% names()
```


## Represent the metaweb

In order to represent this large metaweb, we first represent the network at the SBM group level. We will use then the layout at the aggregated level to compute the layout at the species level. This layout is called `group-TL-tsne`, as `TL-tsne` is the diffusion based layout of `metanetwork`.

### Representation at the SBM level

Using `ggmetanet` and the precomputed layout for `beta = 0.005`, we first represent the web at the SBM group level

```{r SBM level,fig.height = 8, fig.width = 8}
beta = 0.005

#custom ggnet parameters
ggnet.custom = ggnet.default
ggnet.custom$label = T
ggnet.custom$edge.alpha = 0.5
ggnet.custom$alpha = 0.7
ggnet.custom$arrow.size = 1
ggnet.custom$max_size = 12

net_groups = ggmetanet(meta_vrtb,g = meta_vrtb$metaweb_group,flip_coords = T,
          beta = beta,legend = "group",
          ggnet.config = ggnet.custom,edge_thrs = 0.1)
net_groups
```

### group-TL-tsne layout


```{r group-layout,warning=FALSE,fig.height = 8, fig.width = 8}
# beta = 0.005
# ggnet.custom = ggnet.default
# ggnet.custom$label = F
# ggnet.custom$edge.alpha = 0.02
# ggnet.custom$alpha = 0.7
# ggnet.custom$arrow.size = 1
# ggnet.custom$max_size = 4
# ggnet.custom$palette = "Set2"
# #add images in the legend
# ggnet.custom$img_PATH = "vignettes/silouhette_metaweb_europe/"
# 
# 
# net_group_layout = ggmetanet(meta_vrtb,flip_coords = T,mode = "group-TL-tsne",
#                              beta = beta,legend = "group",ggnet.config = ggnet.custom)
# net_group_layout
```



## References

Maiorano, L., Montemaggiori, A., Ficetola, G. F., O’connor, L., & Thuiller, W. (2020). 
TETRA‐EU 1.0: a species‐level trophic metaweb of European tetrapods. Global Ecology and Biogeography, 29(9), 1452-1457.

O’Connor, L. M., Pollock, L. J., Braga, J., Ficetola, G. F., Maiorano, L., Martinez‐Almoyna, C., ... & Thuiller, W. (2020). 
Unveiling the food webs of tetrapods across Europe through the prism of the Eltonian niche. Journal of Biogeography, 47(1), 181-192.

